---
title: "class17 mini project"
author: "faisal"
format: gfm
---
```{r}
# Import vaccination data
vax <- read.csv( "covid19vaccinesbyzipcode_test.csv" )
head(vax)
```

```{r}
vax$as_of_date
```


> Q1. What column details the total number of people fully vaccinated? 

vax$persons_fully_vaccinated
> Q2. What column details the Zip code tabulation area? 

vax$persons_fully_vaccinated

> Q3. What is the earliest date in this dataset? 

```{r}
vax$as_of_date[1]
```



> Q4. What is the latest date in this dataset?

```{r}
vax$as_of_data[nrow(vax)]
```
```{r}
skimr::skim(vax)

```

>Q5. How many numeric columns are in this dataset?

13

>Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column? 

```{r}
sum(is.na(vax$persons_fully_vaccinated))
```


>Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)? 

```{r}
sum(is.na(vax$persons_fully_vaccinated)) / nrow(vax)

```

```{r}
library(lubridate)

```

```{r}
today()

```

We can now magically do math with dates

```{r}
# Specify that we are using the year-month-day format
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
today() - ymd("2021-01-05")
```

```{r}
today() - ymd("2002-06-12")
```

Q9. How many days have passed since the last update of the dataset? 

```{r}
today() - vax$as_of_date[nrow(vax)]
```


Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)? 


```{r}
library(zipcodeR)
geocode_zip('92037')
zip_distance('92037','92109')
reverse_zipcode(c('92037', "92109") )

```

#Focus on the San Diego area


```{r}
# Subset to San Diego county only areas
sd = vax[vax$county == "San Diego" ,]
nrow(sd)
```

```{r}
library(dplyr)

sd <- filter(vax, county == "San Diego")

nrow(sd)
```

```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
```

Q11. How many distinct zip codes are listed for San Diego County? 

```{r}
n_distinct(sd$zip_code_tabulation_area)

```

Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset? 

```{r}
ind <- which.max(sd$age12_plus_population)
sd$zip_code_tabulation_area[which.max(sd$age12_plus_population)]
```
>Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2022-11-15”? 

```{r}
vax$as_of_date[nrow(vax)]
```



```{r}
sd.today <- filter(sd, as_of_date == "2023-02-28")
```

```{r}
mean(sd.today$percent_of_population_fully_vaccinated, na.rm=T)
```

>Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2022-11-15”?

```{r}
hist(sd.today$percent_of_population_fully_vaccinated)
```

```{r}
library(ggplot2)

ggplot(sd.today) +
  aes(percent_of_population_fully_vaccinated) +
  geom_histogram()
```




```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```


> Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
Ucplot <- ggplot(ucsd) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) + 
  ylim(c(0,1)) +
  labs(x= "Date", y="Percentage Vaccinated") 

Ucplot
```

##Comparing to similar sized areas

```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2023-02-28")

#head(vax.36)
```

Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-11-15”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

```{r}
ave <- mean(vax.36$percent_of_population_fully_vaccinated)
ave
```

```{r}
Ucplot + geom_hline(yintercept=ave, col = "red", linetype =2)
```

Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-11-15”?

```{r}
summary(vax.36$percent_of_population_fully_vaccinated)
```

Q18. Using ggplot generate a histogram of this data.

```{r}
ggplot(vax.36) +
  aes(percent_of_population_fully_vaccinated)+
  geom_histogram() 
```

> Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
vax %>% filter(as_of_date == "2022-11-15") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)

```
```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)
```

```{r}
ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  ylim(0,1) +
  labs(x="Date", y= "Percentage Vaccinated",
       title= "Vaccination Rate Across America",
       subtitle="only Areas with a population above 36k are Shown") +
  geom_hline(yintercept = mean(vax.36$percent_of_population_fully_vaccinated), linetype=2)
```

